---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

Abstract;

```{r,echo=FALSE}
f=readRDS("~/Dropbox/chr1500kb.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"500kb.rds"));f=rbind(f,file)}

mvp=f

f2=readRDS("~/Dropbox/chr1500kb_ukbb.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"500kb_ukbb.rds"));f2=rbind(f2,file)}

ukbb=f2
# 
f3=readRDS("~/Dropbox/chr1hdlmin_per_block_derek.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"hdlmin_per_block_derek.rds"));f3=rbind(f3,file)}

f4=readRDS("~/Dropbox/chr1ldlmin_per_block_derek.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"ldlmin_per_block_derek.rds"));f4=rbind(f4,file)}

f5=readRDS("~/Dropbox/chr1tgmin_per_block_derek.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"tgmin_per_block_derek.rds"));f5=rbind(f5,file)}

f6=readRDS("~/Dropbox/chr1tcmin_per_block_derek.rds")
for(i in 2:22){file=readRDS(paste0("~/Dropbox/chr",i,"tcmin_per_block_derek.rds"));f6=rbind(f6,file)}

derek=data.frame(start=f3$V1,stop=f3$V2,hdl=f3$V3,ldl=f4$V3,tg=f5$V3,tc=f6$V3)

colnames(ukbb)=colnames(mvp)=colnames(derek)
```


```{r}
head(derek)
head(ukbb)
head(mvp)
sum(rowSums(mvp[,c(3:6)])>0)
sum(rowSums(derek[,c(3:6)])>0)
sum(rowSums(mvp[,c(3:6)])==0)
sum(rowSums(mvp[,c(3:6)]==0)!=0)

```


We identified XXX novel loci associated with lipids with summary-level data using mashR. FInd the number of nonzero rows in MVP that are zero in derek
```{r}
mvp_list=which(rowSums(mvp[,c(3:6)])>0)
d_list=which(rowSums(derek[,c(3:6)])>0)
length(setdiff(mvp_list,d_list))
```

We increase the total loci linked to lipids by XXX-fold. 

```{r}

sum(rowSums(mvp[,c(3:6)])!=0)/sum(rowSums(derek[,c(3:6)])!=0)
```


In joint replication, XXX of XXX total loci (XXX%) detected in 297,626 MVP participants replicated in XXX UK Biobank participants, 

```{r}
mvp_list=which(rowSums(mvp[,c(3:6)])>0)
ukb_list=which(rowSums(ukbb[,c(3:6)])>0)

length(mvp_list)
length(ukb_list)
length(intersect(mvp_list,ukb_list))
length(intersect(mvp_list,ukb_list))/length(mvp_list)
```

and XXX of XXX total loci (XXX%) detected in UK Biobank replicated in MVP. 

```{r}
length(intersect(mvp_list,ukb_list))/length(ukb_list)
```
