---
title: "SampleSize"
author: "Your Name"
date: "2021-01-21"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

here we want to compare the sample sizes for new versus old:


```{r}
library("data.table")
original.standard.error=fread("~/Downloads/Merged_MVP_Full_se.txt",header = F)
beta.se=as.matrix(data.frame(original.standard.error[,c(3:6)]))

posterior_sd=readRDS("~/Downloads/shared_posteriorSD.rds")
combined_lfsr=readRDS("~/Dropbox/farnambatches/all_MVP_posteriors/combined_lfsr_MVP.rds")
post.var=(posterior_sd*beta.se)^2
```


Now let's plot effective sample size. Recall:

$$n_{jeff}=\frac{s_{j}^2}{\tilde{s_{j}^2}}$$

original sample size here was 297,626 per Klarin et al (klarin et al 2018)

First let's segregate those varaints with significant lfsr:

```{r}
size=297626
ratio=beta.se^2/post.var
njeffective=size*beta.se^2/post.var

colnames(ratio)=colnames(njeffective)=colnames(combined_lfsr)


##clean up
rm(original.standard.error)
rm(posterior_sd)


##divide dataset
sig=which(combined_lfsr<0.05)## identify those with lfsr less than 0.05 in at least one subgroup
not_sig=which(combined_lfsr>0.05)


```              
Recall that in the univariate case, the posterior SD is directly proportional to the prior SD, and so for effects with strongest loading on small $\sigma$ components, the posterior SD will be proportionately smaller (as we are more confident it has mass at prior mean 0).


```{r}
plot(density(unlist(ratio[not_sig])),col="red",main='Density of Ratio')
lines(density(unlist(ratio[sig])),col="green")
```

Correspondingly, the tissues which benefit the most from the sharing boost in precision are those with high standard errors (and correspondingly small Z statistics).

```{r}
barplot(apply(njeffective,2,median),main="Median Effective Sample size")
barplot(apply(ratio,2,median),main="median Increase in Effective Sample Size")
barplot(colSums(combined_lfsr<0.05),main="Number Significant Per Trait")
```

We can consider the posterior variance of $\beta|D$ as $V(\beta/\hat{s}|Data)$*\hat{s}. Recall the univariate result:

$$V(\beta|Data)=\frac{1}{1/\sigma_0^{2}] + 1/\sigma^2}$$ where for the smallest effects, $\sigma_0$ will approach 0, and thus the posterior variance (for us of $$V(beta/\hat{s}|data))$$ will also.

```{r}

plot(density(-log10(post.var[sig])),col="green",main="Density of -log10 V(\beta|data)")
lines(density(-log10(post.var[not_sig])),col="red")
```