---
title: "simualtions"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: no
  html_notebook: default
  html_document:
    df_print: paged
    toc: no
editor_options:
  chunk_output_type: inline
---

# Introduction


Our goal is to simulate GWAS in a situation with 
1) weak signal (only 1% if associations are real)
2) strongly correlated errors (i.e, cor(e_i,e_j)=0.8).

We simulate the data, and then compute posteriors using univariate methods and configuration based joint methods. Then we compute RMSE (measure of accuracy) and produce the ROC curves demonstrating power.

```{r}
setwd("~/lipids_mvp/analysis/")
source("../scripts/Sim_with_error.R")
m.model=readRDS("../data/mashcomplete.rds")
covmat=m.model$fitted_g$Ulist[1:7]
set.seed(123)
sim.data=sim.with.error(J=10000,d=4,rho=0.8,n=100,covmat = covmat,betasd = 0.0001,esd = 0.005) ##sample 100 from the K covariance matrices with equal probability

bmat=sim.data$betahat
semat=sim.data$sebetahat

beta=sim.data$beta

zmat=bmat/semat


```

Here we demonstrate the correlated error matrix:

```{r}
zmash=zmat
library("mashr")
library("flashr")
source('~/Dropbox/jointData/flashscript.R')
# identify a random subset of 20000 tests
random.subset = sample(1:nrow(zmash),2000)

data.temp = mash_set_data(zmash[random.subset,],alpha = 1)
Vhat = estimate_null_correlation_simple(data.temp)
library("lattice")
clrs = colorRampPalette((c("#D73027","#FC8D59","#FEE090","#FFFFBF", "#E0F3F8","#91BFDB","#4575B4")))(64)

print(levelplot(Vhat,col.regions = clrs,xlab = "",ylab = "",colorkey = TRUE,main="VHAT"))

rm(data.temp)
data.random = mash_set_data(zmash[random.subset,],alpha = 1,V=Vhat)

maxval=apply(zmash,1,function(x){max(abs(x))})
zmax=zmash[order(maxval,decreasing = T),][1:100,]


data.strong = mash_set_data(zmax,alpha = 1,V=Vhat)

U.pca = cov_pca(data.strong,3)

#U.flash=cov_flash(data.strong, non_canonical = TRUE)
#X.center = apply(data.strong$Bhat, 2, function(x) x - mean(x))
#U.ed = cov_ed(data.strong, c(U.flash, U.pca, list("XX" = t(X.center) %*% X.center / nrow(X.center))))
#saveRDS(U.ed,"~/lipids_mvp/data/EDcov_sim.Rds")


U.ed=readRDS("~/lipids_mvp/data/EDcov_sim.Rds")
U.c = cov_canonical(data.random)
m = mash(data.random, Ulist = c(U.ed,U.c),outputlevel = 1)

mash.data=mash_set_data(zmash,V = Vhat,alpha = 1)

m$result=mash_compute_posterior_matrices(m, mash.data, algorithm.version = "Rcpp")
saveRDS(m,"~/lipids_mvp/data/mashresult_sim.rds")
```

```{r plotting}

k=length(m$fitted_g$Ulist)
l=length(m$fitted_g$grid)
pimat=matrix(m$fitted_g$pi[-1],nrow=l,byrow=T)
colnames(pimat)=names(m$fitted_g$pi)[2:(k+1)]
barplot(colSums(pimat),las=2)

# library("lattice")
# for(i in 1:k){
#   z.num=as.matrix(cov2cor(m$fitted_g$Ulist[[i]]))
#   colnames(z.num)=row.names(z.num)=colnames(zmash)
# clrs = colorRampPalette((c("#D73027","#FC8D59","#FEE090","#FFFFBF", "#E0F3F8","#91BFDB","#4575B4")))(64)
# z.num[lower.tri(z.num)] = NA
# print(levelplot(z.num,col.regions = clrs,xlab = "",ylab = "",colorkey = TRUE,main=paste0(names(m$fitted_g$Ulist)[[i]])))
# }

library("ashr")

ash.z=apply(zmash,2,function(x){ashr::ash(x,sebetahat = rep(1,length(x)))})
saveRDS(ash.z,"~/lipids_mvp/data/ash_sim.rds")
univariate.ash.pm=cbind(ash.z[[1]]$result$PosteriorMean,ash.z[[2]]$result$PosteriorMean,ash.z[[3]]$result$PosteriorMean,ash.z[[4]]$result$PosteriorMean)

lfsr.ash=cbind(ash.z[[1]]$result$lfsr,ash.z[[2]]$result$lfsr,ash.z[[3]]$result$lfsr,ash.z[[4]]$result$lfsr)


```

If we want to compare to eQTL bma, we use only canonical configurations:

```{r}

U.c = cov_canonical(mash.data)  
print(names(U.c))
```



```{r}
m.c = mash(mash.data, U.c)

bma.pm=m.c$result$PosteriorMean
lfsr.bma=m.c$result$lfsr
```

Now let's do some analysis;

```{r}
sebetahat=semat[1:10000,]
mash.means=m$result$PosteriorMean
lfsr.mash=m$result$lfsr

standard=sqrt(mean((beta[1:10000,]-bmat[1:10000,])^2))
# sqrt(mean((beta[1:10000,]-mash.means[1:10000,]*sebetahat)^2))/standard
# sqrt(mean((beta[1:10000,]-univariate.ash.pm[1:10000,]*sebetahat)^2))/standard
# sqrt(mean((beta[1:10000,]-bma.pm[1:10000,]*sebetahat)^2))/standard

rmse.all.table=cbind(mash=sqrt(mean((beta[1:10000,]-mash.means[1:10000,]*sebetahat)^2))/standard,mash.bmalite=sqrt(mean((beta[1:10000,]-bma.pm[1:10000,]*sebetahat)^2))/standard,ash=sqrt(mean((beta[1:10000,]-univariate.ash.pm[1:10000,]*sebetahat)^2))/standard) 

print("RMSE TABLE",rmse.all.table)

barplot(as.numeric(rmse.all.table),main="Shared, Structured Effects",
        ylab="relative error (RRMSE)",xlab="",col=c("green","blue","red"),names=colnames(rmse.all.table),ylim=c(0,0.15),cex.main=1.5,cex.lab=1,cex.names=1,las=2)
```


```{R}
sign.test.mash=beta*mash.means
sign.test.ash=beta*univariate.ash.pm
sign.test.bma=beta*bma.pm


mash.power=NULL
ash.power=NULL
bma.power=NULL

mash.fp=NULL
ash.fp=NULL
bma.fp=NULL

mash.td=NULL
ash.td=NULL
bma.td=NULL


thresholds=seq(0.01,1,by=0.01)
for(s in 1:length(thresholds)){
thresh=thresholds[s]


##sign power is the proportion of true effects correctly signed at a given threshold
mash.power[s]=sum(sign.test.mash>0&lfsr.mash<=thresh)/sum(beta!=0)
ash.power[s]=sum(sign.test.ash>0&lfsr.ash<=thresh)/sum(beta!=0)
bma.power[s]=sum(sign.test.bma>0&lfsr.bma<=thresh)/sum(beta!=0)

##false positives is the proportion of null effects called at a given threshold
mash.fp[s]=sum(beta==0&lfsr.mash<=thresh)/sum(beta==0)
ash.fp[s]=sum(beta==0&lfsr.ash<=thresh)/sum(beta==0)
bma.fp[s]=sum(beta==0&lfsr.bma<=thresh)/sum(beta==0)

##false sign rate is the proportion of significant effects that are incorrectly signed at a given threshold




}

#frow=c(1,2))

plot(mash.fp,mash.power,cex=0.5,pch=1,col="green",ylab="True Positive Rate",xlab="False Positive Rate",type="l",main="",xlim=c(0,0.4),ylim=c(0,0.6))
#title("True Positive vs False Positive",cex.main=1.5)
lines(ash.fp,ash.power,cex=0.5,pch=1,ylim=c(0,1),col="red")
lines(bma.fp,bma.power,cex=0.5,pch=1,ylim=c(0,1),col="blue")
#lines(eskin.f.fp,eskin.f.power,cex=0.5,pch=1,ylim=c(0,1),col="pink")
#lines(eskin.r.fp,eskin.r.power,cex=0.5,pch=1,ylim=c(0,1),col="purple")
legend("topleft",legend = c("mash","bmalite","ash"),col=c("green","blue","red"),pch=c(1,1,1,1,1))



```
